---
title: "Data Project Documentation"
output: 
  html_document: 
    keep_md: yes
---

### Introduction
The Markdown file documents what was done for the [World Healthcare Data 1995 to 2013 App](https://samyongchang.shinyapps.io/DataProductsProject). It can be found in the hyperlink above.

## Presentation Link
The presenation link for app can also be found [Here](http://rpubs.com/samyongchang/dataproducts).

## Data Source

We will be using the [World Development Indicators](http://data.worldbank.org/topic/health) data base from the World Bank website. The data can be downloaded [here]("http://api.worldbank.org/v2/en/topic/8?downloadformat=csv"). The recode below 

```{r, warning=FALSE, message=FALSE, eval=FALSE}
# Downloading the data

wburl <- "http://api.worldbank.org/v2/en/topic/8?downloadformat=csv"
download.file(wburl, "data.zip")
unzip("data.zip", list=TRUE)
unzip("data.zip", files=c("Metadata_Indicator_8_Topic_en_csv_v2.csv", "8_Topic_en_csv_v2.csv", "Metadata_Country_8_Topic_en_csv_v2.csv", "[Content_Types].xml"))

# Reading the data
d <- read.csv("8_Topic_en_csv_v2.csv", skip=4)
```


# The Code for the App

## Script for ui.R
```{r, eval=FALSE}
# This is for ui.R

library(shiny)
shinyUI(fluidPage(
  
  titlePanel("World Healthcare Data 1995 to 2013"),
  
  sidebarLayout(
    
    sidebarPanel(
      h2("Analyse Country Statistics"),
      selectInput('countryid', "Country", as.character(unique(chartdata$Country.Name))),
      h6("This changes the Summary Statistics for 1995 to 2013 & Comparison with Overall Average tabs"),
      selectInput('stat1', "Choose a Statistic", names(chartdata[,3:15])),
      h6("Select to compare with the overall average in the Comparison with Overall Average tab"),
      width=3
    ),
      
    
    mainPanel(
      
      tabsetPanel(
        position = "above",
        
        tabPanel("Summary Statistics for 1995 to 2013", 
                 verbatimTextOutput("selectedcountry"), 
                 h3("Summary Statistics"),
                 tableOutput("summarytable")), 

        tabPanel("Comparison with Overall Average",
                 h2("Selected Country & Statistics Vs. Overall"),
                 h4("Plot tracks the selected statistic compared with all countries in the dataset.
                    The correlation between the two are also show"),
                 plotOutput("plot2")),          
                
        tabPanel("Motion Plot", 
                 h2("Motion Plot"),
                 h3("Please select the required data in the fields provided"),
                 h4("(Note: Graphics may take a while to load)"),
                 htmlOutput("plot1")),

        tabPanel("Documentation Guide",
                 h6("Summary statistics for 1995 to 2013 - Select a Country and it shows summary data for the country selected"),
                 h6("Comparison with World Average - Select a Country and Statistics and it plots the selected country and the selected statistics compared to the overall"),
                h6("Motion and Comparison Plots - Select the statistics, country that you are interested in to observe how they change over time.")
                 )
          
      )
    )
  )
)
)


```

## Script for server.R
```{r, eval=FALSE}


# Load data from the root directory
chartdata <- readRDS("chartdata.rds")

# Create an table with the average of all attributes
dgroup <- aggregate(chartdata[3:15], by = list(chartdata$Year), FUN="mean")
names(dgroup)[1] <- "Year"

library(shiny)
library(googleVis)
require(devtools)
library(psych)
library(ggplot2)
library(dplyr)


shinyServer(

  
  function(input, output) {
    
  # Summary Tables   
    output$selectedcountry <- renderText({paste("Selected Country:", input$countryid, sep=" ")})
    output$summarytable <- renderTable({
                          describe(chartdata[chartdata$Country.Name==paste(input$countryid,sep=""),])[c(-1,-2),c(3:5,8:10)]
                         })
  # Plot motion chart
    output$plot1 <- renderGvis({
      gvisMotionChart(chartdata,"Country.Name", "Year", 
                           options= list(width=1000, height=500))})
  
  # Plot Line Chart that compares with 
    
    output$plot2 <- renderPlot({
                      
                      num <- match(input$stat1, names(dgroup))
                      
                      cdata <- reactive({chartdata[chartdata$Country.Name==input$countryid,]})
                      charting <- cdata()
                  
                      numc <- match(input$stat1, names(charting))
                      corrnum <- cor(charting[,numc], dgroup[,num])
                      
                      plot(x=charting$Year, y=charting[,numc], type="l", col="steel blue", lwd=4,
                           xlab = "Year", ylab = as.character(input$stat1), ylim=c(min(chartdata[,numc]), max(chartdata[,numc])), xlim=c(1995,2013) 
                           )
                      
                      par(new=T)
                      plot(x=dgroup$Year, dgroup[,num] , type = "l", col="black", xlab="", ylab="", lwd=1, ylim=c(min(chartdata[,numc]), max(chartdata[,numc])), xlim=c(1995,2013))
                      par(new=F)
                      legend("topright", col=c("steel blue","black"), lty=1, lwd=1, bty="n",
                             legend=c(input$countryid, "Overall"))
                      text(1998, max(chartdata[,numc]), paste("Correlation with Overall=", round(corrnum,digits=2), cex=2))
                      
                      })
      
      
  })                            

```

## Scripts for Data Processing

```{r, message=FALSE, warning=FALSE, eval=FALSE}

# Identifying Indicators of Interest
ind1 <- grepl("SH.XPD.OOPC.TO.ZS|
              SH.XPD.OOPC.ZS|SH.XPD.PCAP|SH.XPD.PCAP.PP.KD|
              SH.XPD.PUBL|SH.XPD.PUBL.GX.ZS|SH.XPD.PUBL.ZS|SH.XPD.TOTL.ZS|
              SN.ITK.DEFC.ZS|SP.DYN.CBRT.IN|SP.DYN.CDRT.IN|SP.DYN.LE00.IN|
              SP.POP.0014.TO.ZS|SP.POP.1564.TO.ZS|SP.POP.65UP.TO.ZS|
              SP.POP.GROW|SP.POP.TOTL",d$Indicator.Code)

# We Confine the data to the years 1995 to 2013 where there is Health Expenditure Data for most countries. 
df <- d[ind1,c(-2, -5:-38, -58:-60)]

# Remove Countries that have no data for Health Expenditure, Total (% of GDP)
ind2 <- df[df$Indicator.Name=="Health expenditure, total (% of GDP)",]
ind3 <- as.character(unique(ind2$Country.Name[complete.cases(ind2)]))
df <- df[df$Country.Name %in% ind3,]

# Remove Countries with no data for year 2013
ind4 <- df[is.na(df$X2013),]
df <- df[!df$Country.Name %in% as.character(unique(ind4$Country.Name)),]

# Remove remaining incomplete cases
ind5 <- df[!complete.cases(df),]
df <- df[!df$Country.Name %in% as.character(unique(ind5$Country.Name)),]

# Reshape into Time Series Format
library(reshape2)
library(plyr)
df <- melt(df, id.vars = c("Country.Name", "Indicator.Name", "Indicator.Code"), measure.vars = 4:22)

# Create a New variable for Years
start <- 1994
for (i in unique(df$variable)) {
                          start <- start + 1
                          df$Year[df$variable==paste(i,sep="")] <- start
}


# Create the final data file which consists only of 
selected <- df[,c(1,2,5,6)]
chartdata <- dcast(selected,Country.Name + Year ~ Indicator.Name)
saveRDS(chartdata, file = "chartdata.rds")

```

```{r, echo=FALSE}
chartdata <- readRDS("chartdata.rds")
```

## List of Countries and Indicators in the final Dataset
```{r, results='asis', echo=TRUE}
library(xtable)

# Show Final List of Countries in Data
print(xtable(data.frame(unique(chartdata$Country.Name))), floating=TRUE, comment=FALSE, type = "html", include.rownames = F)

# Shw Final List of Indicators in Data
print(xtable(data.frame(names(chartdata)[3:15])), floating=TRUE, comment=FALSE, type = "html", include.rownames = F)
```


